% tex/sections/discussion.tex
% ──────────────────────────────────────────────────────────────────────────────
\chapter{Discussion}
\label{chap:discussion}
% ──────────────────────────────────────────────────────────────────────────────
%
% TODO: Write discussion.
%
% Suggested structure:
%   - Interpretation of results in light of theory
%   - Comparison with existing literature
%   - Limitations of the study
%   - Alternative explanations


\section{Simple Model}
\label{sec:simple_disc}

\textcolor{magenta}{NO EXPLANATION FOR WHY VX AND STRAIN LOOK SO DIFFERENT AT SNOW BOTTON IN SUB-RAYLEIGH, ALSO EXPLAIN WHY TWO PHASES CAN BE SEEN ONLY IN VY SUPERSHEAR SNOW BOTTOM}


\subsection{Strain and strain rate space-time plots}
The main wavefronts which can be observed in figures \ref{fig:sub_strain} and \ref{fig:super_strain} are the crack fronts. This is because when taking the slope of these two wavefronts, it corresponds to the rupture velocity in the supershear and the sub-rayleigh case. Both cases produce strain and strain rate outputs that look very different, this is important when relating the results here to field measurements. The different behaviors mean that in theory, from the velocity output, the transition point between sub-rayleigh and supershear propagation regimes, as described by \cite{trottetTransitionSubRayleighAnticrack2022,bergfeldSignaturesSubRayleighSupershear2025,sironTheoreticalFrameworkDynamic2023}, could be observed because the strain and strain rate measurement would change. \par

\subsection{Two-wavefront structure in Sub-Rayleigh}
In the sub-rayleigh strain and strain rate plots, two main wavefronts can be observed. One, the main wavefront, which is broadest in the strain rate and strain plots at crack depth in figure \ref{fig:sub_strain}, corresponds in slope to the sub-rayleigh rupture speed, as discussed above. The second wavefront is faster and arrives sooner than the main wavefront and only appears to be emitted from the first source. This changes when source spacing is increased, for example from 0.1m to 10m, as shown in appendix \ref{app:D}, figure \ref{fig:wide_sub_strain}. This leads to the conclusion that the precursor to the main wavefront disappears because of interference. All other sources but the first and the last one have interference from the sources before and after it, which means that this wave type, although emitted, is cancelled out because of interference \parencite{madariagaDynamicsExpandingCircular1976}. When the source spacing is increased, the fronts of the individual sources travel further apart in space and they interfere less. \par

This precursor in the sub-rayleigh strain and strain rate plots has a slope which corresponds to a speed around the p-wave speed in the snow, which is also close to the speed of the symmetric lamb wave mode (S0). The reason for this is is discussed in the velocity section  \ref{sec:disc_sub_two_waves} because it is easier to derive from the velocity output. \par

\subsubsection{Wavefront identification and polarity sequence}

In both the sub-Rayleigh (figure \ref{fig:sub_strain}) and supershear
(figure \ref{fig:super_strain}) cases, all panels show a consistent polaritysequence: positive strain (extension, red) arrives first, followed immediately by negative strain (compression, blue), then a return to positive. This extension-first, compression-second sequence is the expected signature of a compressional P-wave or S0 Lamb mode arrival at a receiver ahead of the source, and is well established in elastic wave propagation theory \parencite{kAkiRichards19801981}. When the crack-tip stress field expands into the undisturbed medium, it first pushes material away from the source (extension), and the elastic restoring force of the medium then pulls it back (compression) before the wavefield equilibrates. The same sequence has been reported in field measurements of avalanches with geophones where p-wave precursors have been observed
\parencite{herwijnenSeismicSensorArray2011}. \par

After this leading extensional arrival comes the dense alternating polarity pattern of the A0 flexural wave train. As derived in section \ref{sec:wave_physics}, the horizontal strain of the A0 mode varies as a cosine in space and time, cycling between maximum tension and maximum compression over each half-wavelength. The detailed arguments of why the A0 Lamb wave mode is present in the results is discussed in the velocity section \ref{sec:vel_disc}. \par

\subsubsection{Dispersion and Broadening of Wavefronts}

In the supershear strain and strain rate plots, the main wavefront weakens with increasing time (meaning the colors are less saturated). The same is true for the precursor in the sub-rayleigh strain and strain rate plots at the snow bottom. The weakening of the wavefront is observed clearly in the supershear case, because the crack travels faster than the guided waves \parencite{dunhamNearsourceGroundMotion2005}. A Mach-front is created, where all the waves travel together in a "shock-front" \parencite{dunhamNearsourceGroundMotion2005}, but once this front passes, there is attenuation of energy. In the sub-rayleigh case on the other hand, the guided waves are faster than the crack \parencite{kAkiRichards19801981}, there is reflection of the waves between the top and bottom of the snow, which creates the osicllations which can be seen in figure \ref{fig:sub_strain}. \par

In the sub-rayleigh case, the precursor wave also broadens and weakens with time (right column of figure \ref{fig:sub_strain}). At the snow bottom, they are further away from their source (which is at crack depth). Because the S0 precursor is only emitted by the first and last source, its first arrival is strongest. After that, it loses energy because there is no source emitting this type of wave anymore, and the S0 precursor loses energy \parencite{kAkiRichards19801981}, which can be seen by the colors fading. This wavefront does not broaden, it only starts fading with increasing time. \par

The broadening of the wavefront with increasing time can be explained by dispersion. By the reflections of the top and bottom snow boundary, the waves present start interfering with each other, which can lead to constructive or destructive interference \parencite{kAkiRichards19801981}. This means that waves can start travelling at different phase velocities, because of the interference \cite{graffWaveMotionElastic1975}. The phase and group velocities of waves are described as \begin{equation}
v_{\text{ph}} = \frac{\omega}{k}, \quad v_{\text{g}} = \frac{d\omega}{dk}
\label{eq:velocities}.
\end{equation} The interference leads to the broadening of the wavefront with increasing time, described by the envelope spreading law \parencite{kAkiRichards19801981} \begin{equation}
\Delta x(t) = \Delta x_0 \sqrt{1 + \left( \frac{\frac{d^2\omega}{dk^2} t}{(\Delta x_0)^2} \right)^2}
\label{eq:envelope_spreading}.
\end{equation} As time goes on, there are more reflections present, so there is more potential for interference and thus dispersion. \par

\subsubsection{Amplitude decay and depth dependence of Strain and Strain Rate}

The weakening of the wavefronts, especially observed in the supershear case in figure \ref{fig:super_strain} and the precursor in \ref{fig:sub_strain} can be explained by depth decay and geometric spreading from the source. The crack tip acts as a moving two-dimensional point source, travelling along a line in a 2D elastic medium, the peak amplitude $A$ of the radiated wavefield decays inversely with the square root of the radial distance $r$ from the source plane \parencite{achenbachWavePropagationElastic2012}:
\begin{equation}
A(r) = \frac{A_0}{\sqrt{r}}
\label{eq:geometric_spreading_2d}
\end{equation}
where $r = \sqrt{(x - x_0)^2 + (y - y_{\text{src}})^2}$. As a consequence, as the wavefield travels vertically through the slab, away from the crack plane, towards the snow bottom, it experiences a geometric reduction in peak energy. \par

Furthermore, for wave components whose horizontal phase velocities are slower than the bulk wave speeds of the snow, the vertical wavenumbers become purely imaginary. Defining $p = i\alpha$ and $q = i\beta$, where $\alpha = \sqrt{k^2 - (\omega/c_p)^2}$ and $\beta = \sqrt{k^2 - (\omega/c_s)^2}$ \parencite{graffWaveMotionElastic1975}, the trigonometric expressions in the wave potentials transform into hyperbolic functions:
\begin{equation}
\phi(x,z,t) = \left[ A \cosh(\alpha z) + B \sinh(\alpha z) \right] e^{i(kx-\omega t)}
\label{eq:hyperbolic_phi}
\end{equation}
\begin{equation}
\psi(x,z,t) = \left[ C \cosh(\beta z) + D \sinh(\beta z) \right] e^{i(kx-\omega t)}
\label{eq:hyperbolic_psi}
\end{equation}
These hyperbolic profiles dictate that high-frequency, shorter wavelength  phases decay exponentially as a function of vertical distance from the location of the crack. In other words, higher frequencies are attenuated more quickly, which explains the weakening of the wavefronts observed. This decay is not dissipative, which means that the energy is not lost, it is trapped in the near-field, which is consistent with the guided-wave trapping described by \textcite{graffWaveMotionElastic1975}. \par



\subsubsection{Magnitude of Wavefronts}

In the supershear case (figure\ref{fig:super_strain}), the two fronts merge into a single broad band because the rupture speed is within a few metres per second of the p-wave speed and the S0 speed. The strain and strain rate panels at both depths look essentially identical with no change in amplitude of strain or strain rate from crack depth to snow bottom. This contrasts sharply with the sub-Rayleigh case \ref{fig:sub_strain}, where the amplitude drops significantly between the two depths.\par

A consequence of this difference in depth dependence of magnitude of strain is its ratio between the two regimes. In the supershear strain rate plots, the colorbar minimum and maximum is approximately one order of magnitude larger than in the sub-Rayleigh panels. Supershear ruptures are more efficient radiators of seismic energy to the free surface because the crack tip is co-located with the wave front rather than trailing behind it, concentrating energy into a Mach cone that hits the
snow surface with maximum constructive interference \parencite{dunhamNearsourceGroundMotion2005, rosakisIntersonicShearCracks2002}.
In the sub-Rayleigh regime, the energy is spread over a broader spatial and
temporal footprint because the rupture front and the precursor waves propagate at different speeds, reducing the peak strain rate at any point in space or time. This amplitude difference has a direct implication for field measurements: supershear avalanche releases would be substantially more detectable by DAS or than sub-Rayleigh events of the same moment release, and consequently, a DAS cable that can just barely detect a supershear event would be unable to detect a sub-Rayleigh event of equal size at the same distance. The same is implied for strain and strain rate at different depths in sub-rayleigh propagation regimes. Because the magnitude of strain (rate) in figure \ref{fig:sub_strain} is one order lower at the snow bottom than at crack depth, the cable location relative to the nucleation is important. Since cables are usually installed before the first snowfall \parencite{edmeFiberopticDetectionSnow2023}, they would be located at the snow bottom rather than at the depth of the crack. This means that the sub-rayleigh crack propagation might be too small in magnitude to be recorded by the DAS cable. Since \cite{bergfeldSignaturesSubRayleighSupershear2025,sironTheoreticalFrameworkDynamic2023,trottetTransitionSubRayleighAnticrack2022, gaumeDynamicAnticrackPropagation2018} have shown that there can be a transition between sub-rayleigh and supershear crack propagation, this means that the sub-rayleigh onset of the crack might not be recorded in DAS because of its smaller magnitude. \par



% The sub-rayleigh horizontal strain and strain rate plots in figure \ref{fig:sub_strain} look very similar to the velocity plots in figure \ref{fig:sub_vel}. The same is true for the supershear strain and strain rate in figure \ref{fig:super_strain}, which look similar to all but the vertical velocity at the snow bottom (top right in figure \ref{fig:super_vel}). This is intuitive because strain is the spatial derivative of displacement ($\ \frac{\partial u}{\partial x}$), strain rate is the temporal derivative of strain ($\ \frac{\partial^2 u}{\partial x \partial t}$). Velocity on the other hand is the spatial derivative of displacement ($\ \frac{\partial u}{\partial t}$), so all the values are derived from displacement. The first wave caused by lack of interference \parencite{madariagaHighfrequencyRadiationCrack1977} in the sub-rayleigh case can also be observed in the strain and strain rate plots, this disappears again when source spacing is increased (appendix \ref{app:D}, figure \ref{fig:wide_sub_strain}). \par


% Both the sub-rayleigh and the supershear strain and strain rate cases show a wave train running ahead. The slope of the first wavefront observed in figures \ref{fig:sub_strain,fig:super_strain} is close to the p-wave velocity in snow and the rupture velocity in the supershear case, which as discussed, is also close to the velocity of the S0 mode. This means that in both cases, the S0 wave modes are the first arrivals because they are fastest, likely followed by the p-waves. This is consistent with the observations made in the velocity plots. In the supershear case, the S0 waves are followed immediately by the waves from the rupture front, and the p-waves which would explain why there is only one clear but very broad wavefront in figure \ref{fig:super_strain}. In the sub-rayleigh strain and strain rate plots in figure \ref{fig:sub_strain}, the observations are also consistent with velocity. The first wave front to arrive are the p-waves and the S0 mode, followed by the front from the rupture and the s-waves last. The same goes for the widening of the wavefronts, which has already been discussed in the velocity section. \par 

% The first arrival of the p-wave and S0 mode is also consistent with wave propagation. All plots in figures \ref{fig:sub_strain,fig:super_strain} show positive horizontal strain or strain rate first, followed by compression (negative strain) and positive strain again. This propagation is characterized by extension first and then contraction \parencite{kAkiRichards19801981}. \par

% This very repetitive sequence of opposite polarities is a consequence of the snow acting as a waveguide, as described above. When energy is suddenly released by a propagating crack tip, the crack acts as an impulsive source that projects disturbances into the overlying slab \parencite{heierliAnticrackNucleationTriggering2008, gaumeDynamicAnticrackPropagation2018}. The leading P-wave arrival corresponds to the high-frequency, long-wavelength limit of the symmetric ($S_0$) Lamb mode, which represents an in-plane, extensional push-pull wave packet \parencite{achenbachWavePropagationElastic2012, graffWaveMotionElastic1975}. As this transient $S_0$ pulse outruns the rupture front, the initial forward displacement stretches the material, yielding a leading zone of positive tensile strain (red color). To satisfy elastodynamic restoring forces and conserve momentum as the wavefront passes a fixed point in space, this  extension must be immediately followed by a compensating contractional phase of negative compressive strain (blue), before returning to equilibrium. \par 

% Behind this fast, extensional precursor lies the slower, highly oscillatory wave train, which displays a dense pattern of alternating opposite polarities. This trailing pattern is dictated by the superposition of the symmetric mode as well as the antisymmetric ($A_0$) flexural Lamb wave mode. \par

% To relate the Lamb wave modes derived in section \ref{sec:wave_physics}to strain, the horizontal displacement $u_x$ is computed from the potentials:
% \begin{equation}
% u_x = \frac{\partial \phi}{\partial x} - \frac{\partial \psi}{\partial z} = \left[ ik B\sin(pz) + q C\sin(qz) \right] e^{i(kx-\omega t)}
% \label{eq:ux_asym}
% \end{equation}

% Taking the spatial derivative with respect to $x$ to find the horizontal strain ($\varepsilon_{xx} = \partial u_x / \partial x$) transforms the expression via multiplication by $ik$ in the wavenumber domain:
% \begin{equation}
% \varepsilon_{xx}^{\text{asym}}(x, z, t) = \left[ -k^2 B\sin(pz) + ikq C\sin(qz) \right] e^{i(kx-\omega t)}
% \label{eq:strain_asym_final}
% \end{equation}

% By isolating the real part of this analytical expression, the spatial and temporal variation of the flexural strain field isolates the reason for the opposite polarity patterns along the wave train. The strain can be rewritten in terms of its harmonic phase:
% \begin{equation}
% \varepsilon_{xx}^{\text{asym}}(x, z, t) = \mathcal{A}_{\text{asym}}(z) \cos(kx - \omega t + \phi_0)
% \label{eq:harmonic_strain_real}
% \end{equation}
% where $\mathcal{A}_{\text{asym}}(z)$ represents the depth-dependent amplitude envelope and $\phi_0$ is a constant phase angle. The transcendental cosine term dictates that over a fixed spatial distance of exactly half a wavelength, $\Delta x = \lambda/2 = \pi/k$, the wavefield must cycle smoothly between its maximum positive value ($\cos = +1$) and its maximum negative value ($\cos = -1$). 

% Physically, this represents the dynamic bending or structural "ringing" of the slab  \parencite{heierliAnticrackNucleationTriggering2008, bergfeldSignaturesSubRayleighSupershear2025}. When a section of the slab flexes vertically, the fibers located on a single side of the neutral axis experience localized horizontal squeezing ($\varepsilon_{xx} < 0$, blue bands). As the elastic wave oscillates into its next half-cycle, the local curvature reverses, subjecting those same fibers to intense horizontal stretching ($\varepsilon_{xx} > 0$, red bands). 

% When this is translated to the horizontal strain rate ($\dot{\varepsilon}_{xx}$), the time derivative ($\partial/\partial t \longleftrightarrow -i\omega$) shifts the wave profile into a sine function:
% \begin{equation}
% \dot{\varepsilon}_{xx}^{\text{asym}}(x, z, t) = \omega \mathcal{A}_{\text{asym}}(z) \sin(kx - \omega t + \phi_0)
% \label{eq:harmonic_rate_real}
% \end{equation}
% This differentiation scales the amplitude by the angular frequency $\omega$ and introduces a $90^\circ$ spatial phase shift, but leaves the underlying alternating pattern of opposite polarities completely intact. Because both the crack depth ($y = 2.625\,\text{m}$) and the snowpack base ($y = 2.25\,\text{m}$) reside entirely within the lower half of the slab waveguide below its neutral axis ($z < 0$), they evaluate to the same geometric sign in $\sin(pz)$. Consequently, both recording horizons register these rapid, cyclic transitions from extension to contraction perfectly in-phase with one another. 


\subsection{Velocity space-time plots}
\label{sec:vel_disc}
Like in the strain and strain rate plots, the main wavefronts in the supershear and sub-rayleigh case in figures \ref{fig:super_vel,fig:sub_vel} correspond in slope to the rupture speed of each case. Like in strain (rate), the velocity output is vastly different for sub-rayleigh and supershear, which means a transition between propagation regimes can not only be observed in the strain output, but also in the velocity outupt. \par

\subsubsection{Similarity to Strain and Strain Rate Plots}
The sub-rayleigh horizontal strain and strain rate plots in figure \ref{fig:sub_strain} look very similar to the horizontal velocity plots in figure \ref{fig:sub_vel}. The same is true for the supershear strain and strain rate in figure \ref{fig:super_strain}, which look similar to all but the vertical velocity at the snow bottom (top right in figure \ref{fig:super_vel}). This is intuitive because strain is the spatial derivative of displacement ($\ \frac{\partial u}{\partial x}$), strain rate is the temporal derivative of strain ($\ \frac{\partial^2 u}{\partial x \partial t}$). Velocity on the other hand is the spatial derivative of displacement ($\ \frac{\partial u}{\partial t}$), so all the values are derived from displacement. The strain and strain rate shown is in the horizontal direction, which explains why the vertical velocity looks different from the strain (rate) plots. In this section, only results which are specific to velocity will be discussed in detail, all other results are discussed and proved in detail in the strain and strain rate discussion. \par

Because the velocity plots are so similar to the strain and strain rate plots, the effects observed in both results are similar. Just like in strain and strain rate, there is broadening of the wavefront in the supershear case in figure \ref{fig:super_vel}, which can be explained by dispersion and envelope spreading, shown in the strain and strain rate discussion. \par

The amplitude decay of the main wavefront and the precursor wave discussed in the strain rate plots is present in the supershear case and the sub-rayleigh vx at crack depth, it can also be explained by amplitude decay and depth dependence. \par

\subsubsection{Two-wavefront structure in Sub-Rayleigh}
\label{sec:disc_sub_two_waves}

The precursor wave which can be observed in the vx plots in the sub-rayleigh case in figure \ref{fig:sub_vel} has a much faster speed than the rupture itself. This wave also only occurs in the sub-rayleigh case in the vx plots. As for strain, simulations with larger source spacing (for example 10m spacing instead of 0.1m spacing, see appendix \ref{app:D}, figure \ref{fig:wide_sub_vel}) have shown that this precursor disappears if source spacing is increased. This leads to the same conclusion as discussed in the strain case that this is because of interference between the sources \parencite{madariagaDynamicsExpandingCircular1976}. \par

The precursor wavefront is the symmetric Lamb wave mode (S0), whose propagation speed is limited by the plate wave speed: \begin{equation}
  c_p^{plate} = 2c_s\sqrt{1-\frac{v_s^2}{v_p^2}}
\end{equation} \parencite{roseWavesPlates2014}. When using the material parameters used in the simulation, this is just short of the p-wave speed of the slab. The speed of the S0 wave converges to the plate wave speed for low frequencies (long wavelength limit) \parencite{roseWavesPlates2014}, which is why this wave is believed to be the S0 mode which is excited. \par

This two-wavefront structure with a fast S0/P-wave precursor followed by a slower crack-front signal is the signature of sub-Rayleigh rupture. The
spatial separation between the two fronts grows linearly in time as
\begin{equation}
  \Delta x(t) = (v_p - v_r)\,t,
  \label{eq:front_separation}
\end{equation} where \(v_p\) is the P-wave speed and \(v_r\) is the rupture speed. This linear growth is directly measurable from the space-time plots in figure \ref{fig:sub_vel} and is consistent with the theoretical prediction for an expanding crack in a homogeneous elastic medium \parencite{freundDynamicFractureMechanics1998}. The origin of the precursor is the redistribution of the stress field ahead of the crack tip, which propagates at the fastest available wave
speed in the medium regardless of the crack tip speed
\parencite{kAkiRichards19801981}, so in this case, the p-wave or the S0 wave speed.Because the p-wave speed is very close to the S0 speed, this result is consistent with field observations of avalanches using seismic geophone arrays by \cite{herwijnenSeismicSensorArray2011}, a p-wave precursor to the main avalanche release was also observed. While \cite{herwijnenSeismicSensorArray2011} do not make note of the crack propagation regime, the p-wave or S0 precursor can be an important warning signal for sub-rayleigh crack propagation leading to avalanche release. \par

The precursor is visible in sub-Rayleigh because the separation $\Delta x(t)$ grows fast enough to resolve the two fronts within the observation window. In the supershear case, the rupture front, S0 and the pressure wave all have velocities within a couple of metres per second of each other. Nevertheless, in the vertical velocity at the snow bottom in  \ref{fig:super_vel}, two more distinct wavefronts can be seen, one which could be the S0 wave. In all supershear velocity plots, at the end of the crack, there is a fanning out of waves which can be observed, which could be a combination of the stopping phases of the different wave types. This behaviour at arrest is consistent with the stopping-phase radiation described by \textcite{madariagaHighfrequencyRadiationCrack1977}, who showed that abrupt arrest of a crack generates a burst of high-frequency energy that propagates at the bulk wave speeds of the medium, so the p-wave speed, which is close to the S0 in this case, independent of the rupture speed before the arrest. \par

\subsubsection{Lamb Wave Identification}

The presence of waves with opposite polarity can be proven mathematically, the detailed derivation of this is shown in section \ref{sec:wave_physics}.
  
The animations of the velocity wavefield in appendix \ref{app:F} show the propagation intensity of the velocity wavefield. When comparing the horizontal velocity vx in figure \ref{fig:vx_animated_wavefield} with the wavefield intensity of the two lamb wave modes, for example in \cite{cartaLambWavesDiscrete2023}, which can be found in appendix \ref{fig:lamb_wave_modes}, it becomes evident that the main mode which can be observed is the antisymmetric (A0) mode. This is because at source depth, the velocity field has a lower intensity. Above and below the sources, there are narrower bands of low intensity (blue in figure \ref{fig:vx_moving_crack}, red in the reference figure \ref{fig:lamb_wave_modes}) beside bands of higher intensity. This can be explained by the fundamental theory of lamb waves \parencite{lambWavesElasticPlate1917,roseWavesPlates2014}. Because the moment tensor used here has a distinct out of plane component (negative myy), the primary mode which is excited is A0 \parencite{roseWavesPlates2014}. As discussed, the S0 mode can also be seen, but is a precursor rather than the main wavefront and carries less energy. This precursor can also be seen in figure \ref{fig:vx_moving_crack}. \par 

The dominance of A0 over S0 follows from source-mode coupling theory. For a moment tensor source applied normal to the slab midplane (so normal to the horizontal of the slab), the coupling coefficient to A0 scales as the bending moment integrated over the source region, while coupling to S0 scales as the in-plane force resultant \parencite{roseWavesPlates2014, achenbachWavePropagationElastic2012}. Because the force applied has a negative y-component (Myy of the moment tensor) and is normal to the slab axis, this maximizes A0 excitation. Because the force is only in the negative direction, it is asymmetric, so while A0 excitation is maximized, S0 coupling is minimized at low \(f \cdot h\). This theoretical expectation is confirmed by the simulations: S0 appears primarily as a precursor in \(v_x\) (in-plane, extensional motion), consistent
with the mode shapes described by \textcite{lambWavesElasticPlate1917} and the coupling analysis of \textcite{roseWavesPlates2014}. This identification is strengthened by the animation stills in appendix \ref{app:F}, especially vx in figure \ref{fig:vx_moving_crack}. The velocity field shows a polarity reversal around the crack, meaning that for example, the panel above the crack has positive polarity whereas the one below has negative polarity. Alongside of the comparison with figure \ref{fig:lamb_wave_modes} in appendix \ref{app:G}, this is proof for the A0 mode because S0 would have the same polarities on either side of the crack \parencite{lambWavesElasticPlate1917}. \par



\subsection{Discussion of fk-plot}

Similarly to the results for the velocity and strain plots, the fk-plots shown in section \ref{sec:fk_results} show that the energy is concentrated around the p-wave speed (red line in figure \ref{fig:preliminary_fk}) and the rupture speed (blue line in figure \ref{fig:preliminary_fk}). In the supershear case, these two speeds are very close together, which leads to one energy band, whereas in the sub-rayleigh case, they are more separated and two energy bands are observed. The S0 wave speed, as discusse previously, is also very close to the p-wave speed, which also has an effect on the concentration of energy n those areas. \par

In the sub-Rayleigh f-k plot in figure~\ref{fig:preliminary_fk}, the energy
between the two dominant slopes (between the rupture speed and the P-wave speed lines) is not random noise. It corresponds to the dispersive
A0 Lamb mode, which at different frequencies propagates at different phase
velocities between \(v_r\) and \(v_p\). The fan of diagonal stripes observed in the sub-Rayleigh \(v_y\) waterfall plots maps onto this wedge in f-k space: each stripe is a different frequency component of A0 arriving at a different time at a given location. This dispersive fan is a direct encoding of the A0 dispersion curve of the snow slab at \(f \cdot h \approx 6\) Hz\,m \parencite{roseWavesPlates2014, lambWavesElasticPlate1917}. In principle, the shape of this dispersive wedge could be inverted to recover the elastic properties of the slab (thickness, shear modulus) using the same surface-wave dispersion
inversion methods applied in geophysical prospecting \parencite{yilmazSeismicDataAnalysis2001}. \par

In the sub-Rayleigh case, the f-k plot in figure \ref{fig:preliminary_fk} shows a concentration of energy at the \(v_p\) of the slab and the S0 mode, as well as the rupture speed, consistent with the observations made in the velocity and strain plots. \par

This is also the case in the supershear f-k plots and the corresponding velocity and strain plots. Because the rupture speed is very close to the P-wave and S0 speed in the medium, energy is concentrated around these two. The narrowness of the energy concentration in the supershear case compared to sub-Rayleigh reflects the absence of the dispersive two-front fan: the Mach cone geometry radiates energy preferentially at the cone angle \(\xi = \sin^{-1}(c_s/v_r)\) \parencite{dunhamNearsourceGroundMotion2005, rosakisIntersonicShearCracks2002},
which at \(v_r \approx v_p\) collapses to a nearly flat front, concentrating all radiated energy into a single, tightly bounded slope in f-k space. \par

The strong contrast in how the energy is concentrated between the sub-rayleigh and supershear case in \ref{fig:preliminary_fk} can be explained by the number of locations the energy is concetrated. As described above, the supershear regime has energy at essentially
one slope, whereas the sub-Rayleigh case fills a wedge between two slopes. This explains why the energy bands in the supershear case appear much narrower and more concentrated compared to sub-Rayleigh. \par 

The observation that at higher frequencies in figure \ref{fig:preliminary_fk}, the lines of the energy present do not match as well anymore with the concentration of energy, can be explained by numerical dispersion. As \cite{komatitschIntroductionSpectralElement1999} describe, as frequency increases, the wavelength decreases. This means that higher frequencies are not resloved as well by a spatial grid with a certain spacing \cite{yilmazSeismicDataAnalysis2001}, this starts bending the energy in the fk plot. This could explain the mismatch between the physical velocity lines and the concetrations of energy. \par

The horizontal striping visible in the sub-Rayleigh f-k plot at constant frequency (for example radiating from the vp-line) is a temporal aliasing artefact arising from the finite output sampling interval of the simulation, not a physical feature. This can be distinguished from the physical dispersion fan by the fact that the stripes are perfectly horizontal (constant frequency, for all wavenumbers), whereas physical A0 dispersion produces curved or oblique features in f-k space. \par

\subsection{Implications for DAS measurements \& Synthesis of Results}

The results in section \ref{sec:results-1} and their discussion have shown that there is multiple implications for the measurement and interpretation of DAS data, even in a simple physics case. \par 

In all sub-rayleigh results, the presence of a precursor wavefront, which is the S0 wave, is shown. This is consistent with field observations made by \cite{herwijnenSeismicSensorArray2011}, but the precursor is also weaker in amplitude compared to the main wavefront and decays in amplitude with time. This is important because it shows that the precursor wave might not be a relaible "early warning" because it might not be detected on account of its low magnitude. The main wavefront travels at crack propagation speed and does not undergo significant amplitude decay. \par

In all supershear results, there is only one wavefront because the S0 and p-waves have a similar speed to the rupture. The precursor can be identified by looking at the vertical velocity at the snow bottom as well as the rupture arrest. In the supershear case, the wavefront is dispersive and decays in amplitude with depth, it also broadens with increasing propagation time. This is because the rupture and the precursors travel together as a Mach front, which is attenuated \parencite{dunhamNearsourceGroundMotion2005}. The broadeining can be explained by dispersion, so the change in phase velocities caused by reflections and interference \parencite{graffWaveMotionElastic1975}. \par

The supershear velocities and strain (rate) has a higher amplitude than their sub-rayleigh counterparts. This means that supershear ruptures can be detected more easily in DAS measurements because they radiate more energy. In the sub-rayleigh case, there is also significant difference in the amplitudes of strain, strain rate and velocity between the crack depth and the bottom of the snow, the bottom of the snow having tenfold lower amplitudes. This is crucial for DAS field measurements because the cable is usually located at the bottom of the snow. The implications for field measurements are that the sub-rayleigh propagation preceding the supershear transition \parencite{trottetTransitionSubRayleighAnticrack2022} might not be seen in DAS data because it has too low amplitude. This is important when interpreting field data because false conclusions of pure supershear rupture might be drawn from this. \par 



\section{Material Point Method Guided Simulations}
\label{sec:mpm_disc}

\subsection{MPM strain and strain rate plots}

\subsection{MPM fk-plots}

\section{Comparison between pure SE and MPM}
