% tex/sections/outlook.tex
% ──────────────────────────────────────────────────────────────────────────────
\chapter{Outlook}
\label{chap:outlook}
% ──────────────────────────────────────────────────────────────────────────────
%
% TODO: Write outlook.
%
% Suggested structure:
%   - Future work directly motivated by this thesis
%   - Broader research directions opened up
%   - Proposed experiments or improvements to methodology

Based on the results and methods presented in this thesis, there are many open lines of inquiry. Because the model described in the simple case is a basic two layer model, a next step could be to add multiple layers, especially because cracks in snow usually propagate in a weak layer \parencite{schweizerSnowAvalancheFormation2003}. Here, there is also possibility to run the simulations described in this thesis for different snow parameters (density, vp and vs, elastic modulus), because \cite{barteltPhysicalSNOWPACKModel2002,mulakNumericalInvestigationMixedmode2019,bobillierMicromechanicalModelingSnow2020} have all shown that snow properties not only change over time, but also have an influence on crack propagation. This would be useful to better understand DAS data from different times of the year or day, when there is the possibility of snow properties having changed. \par

There is also the possibility of adding topography or a slope angle to the simulation, because as indicated by \cite{anceySnowAvalanches2001,trottetTransitionSubRayleighAnticrack2022}, avalanches are gravity driven and the crack propagation regimes can change greatly with the presence of topography. \par 

Because of the possibility for transition between the different propagation regimes, meaning that the propagation speed changes, as described by \cite{trottetTransitionSubRayleighAnticrack2022}, which is not only driven by topography and slope angle. Another next step would be to have source propagation accelerate with increasing propagation distance, instead of having a constant propagation speed, which is not very plausible on a mountain slope. \par 

Furthermore, the crack modeled in this thesis is linear and at constant depth. MPM simulations of weak layer cracking and collapse have shown that this is not the case \parencite{trottetTransitionSubRayleighAnticrack2022}, so a next step would be to make the crack propagation modeled by seismic sources more physical. This could be done by having "jumps" of the crack into different depths of the snowpack or having multiple lines of sources firing at the same time to represent multiple smaller cracks. \par

The last step in making the model described here physically accurate is to implement the 2D model described here in 3D, which would help to understand slope DAS data even better. Such a 3D model could represent data recorded further away from the crack, in this thesis, it is always assumed the cable is buried in the crack or right below it. It would also be very interesting to compare different cable orientations relative to the crack to see if the wavefield data recorded changes. \par 

